---
title: "Untitled"
output: pdf_document
date: "2024-12-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Make sure to install packages before using them

```{r}
library(tidyverse)
library(plyr)
library(lme4)
library(plotfunctions)
library(dplyr)
library(Amelia)
library(tidyr)
library(corrplot)
library(ggplot2)
library(performance)
```

```{r}
setwd("C:/Users/Thanos/Desktop/advanced statistical modeling/mental health")
```
```{r}
rstudioapi::filesPaneNavigate("C:/Users/Thanos/Desktop/advanced statistical modeling/mental health")

```

```{r}
#read.csv("students_mental_health_survey.csv")
data <-read.csv("students_mental_health_survey.csv")

```

```{r}
str(data)
summary(data)
```

So the data table has very few missing values, however because later on I encountered some error due to the uneven length of the collumns I decided to weed the records with missing values out.

```{r}
col_na_count <- colSums(is.na(data))
print(col_na_count, max = length(col_na_count))
```
Don't bother with the missmap, it turns out I was trying to find something that isn't even there, I believe the NA values were so few that missmap function can't show much detail
```{r}
missmap(data, main = "Missing Values Map", col = c("red", "blue"))

```
```{r}

data %>%
  summarise(across(where(is.numeric), list(
    mean = ~mean(.x, na.rm = TRUE),
    median = ~median(.x, na.rm = TRUE),
    sd = ~sd(.x, na.rm = TRUE)
  )))
```
```{r}
mean_table <- data %>%
  summarise(across(where(is.numeric), ~mean(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Mean")

median_table <- data %>%
  summarise(across(where(is.numeric), ~median(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Median")

sd_table <- data %>%
  summarise(across(where(is.numeric), ~sd(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "SD")

```


```{r}
# Correlation matrix
cor_matrix <- data %>%
  select(where(is.numeric)) %>%
  cor(use = "complete.obs")

# Visualize the correlation matrix
corrplot(cor_matrix, method = "circle", type = "lower", tl.cex = 0.8)

```
That's where I omit NA values
```{r}
# NOT NEEDED, THE MEESING VALUES ARE TOO FEW
clean_health <- na.omit(data[, c("Age","Course", "Gender", "CGPA", "Stress_Level", "Depression_Score", "Anxiety_Score", "Sleep_Quality", "Physical_Activity", "Diet_Quality","Social_Support", "Relationship_Status", "Substance_Use", "Counseling_Service_Use", "Family_History", "Chronic_Illness", "Financial_Stress", "Extracurricular_Involvement", "Semester_Credit_Load", "Residence_Type")])
```

Really just used my intuiton to write down some linear models that made some sense heuristicaly 
```{r}
model1 <- lm(Depression_Score ~ Age + Gender, clean_health)
model2 <- lm(Depression_Score ~ Course + CGPA, clean_health)
model3 <- lm(Depression_Score ~ CGPA, clean_health)
model4 <- lm(Depression_Score ~ Stress_Level, clean_health)
model5 <- lm(Depression_Score ~ Anxiety_Score, clean_health)
model6 <- lm(Depression_Score ~ Sleep_Quality + Physical_Activity + Diet_Quality, clean_health)
model7 <- lm(Depression_Score ~ Social_Support, clean_health)
model8 <- lm(Depression_Score ~ Relationship_Status, clean_health)
model9 <- lm(Depression_Score ~ Substance_Use, clean_health)
model10 <- lm(Depression_Score ~ Counseling_Service_Use, clean_health)
model11 <- lm(Depression_Score ~ Chronic_Illness, clean_health)
model12 <- lm(Depression_Score ~ Financial_Stress, clean_health)
model13 <- lm(Depression_Score ~ Extracurricular_Involvement, clean_health)
model14 <- lm(Depression_Score ~ Semester_Credit_Load, clean_health)
```

```{r}
plot(model1)
plot(model2)
plot(residuals(model2))
plot(model2, which = 1)
plot(model2, which = 2)
qqnorm(residuals(model1))
qqline(residuals(model1))
qqnorm(residuals(model2))
qqline(residuals(model2))
```


```{r}
#the following plot shows that the Depression Score is getting smaller the higher the GPA becomes, which is unexpected and interesting.
plot1 <-ggplot(clean_health, aes(x = CGPA, y = Depression_Score)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(title = "Depression Score vs. CGPA", x = "CGPA", y = "Depression Score") +
  theme_minimal()
```

```{r}
plot1
```
```{r}
ggsave(plot1, 
       filename = "cgpa_vs_depression.png",
       device = "png",
       height = 3, width = 6, units = "in")
```

```{r}
#Doing the same for each course shows a similar trend, you can also see that the Computer Science students have a significantly larger Depression score
plot2 <-ggplot(clean_health, aes(x = CGPA, y = Depression_Score, color = Course)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Depression Score vs. CGPA by Course", x = "CGPA", y = "Depression Score") +
  theme_minimal()
```
```{r}
plot2
ggsave(plot2, 
       filename = "cgpa_vs_depression_each.png",
       device = "png",
       height = 3, width = 6, units = "in")
```


```{r}
#On the relation between stress level and depression score, there is slight trend that shows that the stress level has an negative relation with depression score, which I found weird.
plot3 <-ggplot(clean_health, aes(x = Stress_Level, y = Depression_Score, color = Course)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Depression score vs. Stress level", x = "Stress level", y = "Depression Score") +
  theme_minimal()
```

```{r}
plot3
ggsave(plot3, 
       filename = "stress_vs_depression.png",
       device = "png",
       height = 3, width = 6, units = "in")
```

```{r}
#On the following plot you can observe a slight possitive relation between credit load and depression score, although as you can see except from the Computer science students all the others are grouped up very closely together
plot4 <-ggplot(clean_health, aes(x = Semester_Credit_Load, y = Depression_Score, color = Course)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Depression score vs. Credit load", x = "Credit load", y = "Depression Score") +
  theme_minimal()
```

```{r}
plot4
ggsave(plot4, 
       filename = "credit_load_vs_depression_each.png",
       device = "png",
       height = 3, width = 6, units = "in")
```

```{r}
#Also very unexpected, the Relationship status attribute has little to no effect at least on its own to the depression score.
ggplot(clean_health, aes(x = Relationship_Status, y = Depression_Score)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Depression Score by Relationship Status",
       x = "Relationship Status", 
       y = "Depression Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
The idea was to calculate the mean Depression_Score for each non numerical feature. I was hopping to see significant difference for example in mean Depression Score for males and females but in general the only category that showed significant difference was the Computer Science category in Course, which is much higher that the rest. You can also see that on the following plot
```{r}
plot6 <-ggplot(clean_health, aes(x = Course, y = Depression_Score)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Depression Score by Course", x = "Course", y = "Depression Score") +
  theme_minimal()
```

```{r}
plot6
ggsave(plot6, 
       filename = "course_depression.png",
       device = "png",
       height = 3, width = 6, units = "in")
```

```{r}
mean_depression_course <- dplyr::group_by(clean_health, Course) %>%
    dplyr::summarize(Mean_Depression_Score = mean(Depression_Score, na.rm = TRUE)) %>%
    dplyr::ungroup()

print(mean_depression_course, n = Inf)

```

```{r}

# Calculate the mean Depression_Score for each non-numerical feature


mean_depression_gender <- dplyr::group_by(clean_health, Gender) %>%
  dplyr::summarise(Mean_Depression_Score = mean(Depression_Score, na.rm = TRUE))

mean_depression_sleep <- dplyr::group_by(clean_health,Sleep_Quality) %>%
  dplyr::summarise(Mean_Depression_Score = mean(Depression_Score, na.rm = TRUE))

mean_depression_physical <- dplyr::group_by(clean_health, Physical_Activity) %>%
  dplyr::summarise(Mean_Depression_Score = mean(Depression_Score, na.rm = TRUE))

mean_depression_diet <- dplyr::group_by(clean_health, Diet_Quality) %>%
  dplyr::summarise(Mean_Depression_Score = mean(Depression_Score, na.rm = TRUE))

mean_depression_social <- dplyr::group_by(clean_health, Social_Support) %>%
  dplyr::summarise(Mean_Depression_Score = mean(Depression_Score, na.rm = TRUE))

mean_depression_relationship <- dplyr::group_by(clean_health, Relationship_Status) %>%
  dplyr::summarise(Mean_Depression_Score = mean(Depression_Score, na.rm = TRUE))

mean_depression_substance <- dplyr::group_by(clean_health, Substance_Use) %>%
  dplyr::summarise(Mean_Depression_Score = mean(Depression_Score, na.rm = TRUE))

mean_depression_counceling <- dplyr::group_by(clean_health, Counseling_Service_Use) %>%  dplyr::summarise(Mean_Depression_Score = mean(Depression_Score, na.rm = TRUE))

mean_depression_physical <- dplyr::group_by(clean_health, Physical_Activity) %>%
  dplyr::summarise(Mean_Depression_Score = mean(Depression_Score, na.rm = TRUE))

mean_depression_family <- dplyr::group_by(clean_health, Family_History) %>%
  dplyr::summarise(Mean_Depression_Score = mean(Depression_Score, na.rm = TRUE))

mean_depression_chronic <- dplyr::group_by(clean_health, Chronic_Illness) %>%
  dplyr::summarise(Mean_Depression_Score = mean(Depression_Score, na.rm = TRUE))

mean_depression_extra <- dplyr::group_by(clean_health, Extracurricular_Involvement) %>%
  dplyr::summarise(Mean_Depression_Score = mean(Depression_Score, na.rm = TRUE))

mean_depression_residence <- dplyr::group_by(clean_health, Residence_Type) %>%
  dplyr::summarise(Mean_Depression_Score = mean(Depression_Score, na.rm = TRUE))

```

Here I just performed varius anovas, the model2 that contains the Course is alot better than any other, in general I believe that the "Course" feature will account for alot of information compared to the others
```{r}
anova(model1,model2)
anova(model2,model3)
anova(model3,model4)
anova(model4,model5)
anova(model5,model6)
anova(model6,model7)
anova(model7,model8)
anova(model8,model9)
anova(model9,model10)
anova(model10,model11)
anova(model11,model12)
anova(model12,model13)
anova(model13,model14)
```
Likewise I wrote down some more complex random effects models that made some sence to me
```{r}

model <- lmer(Depression_Score ~ Stress_Level + Anxiety_Score + Sleep_Quality + Financial_Stress +
                Social_Support + Diet_Quality + (1 | Course), data = clean_health,REML=FALSE)

summary(model)

modell <- lmer(Depression_Score ~ Stress_Level + Anxiety_Score  + Financial_Stress +
                  Diet_Quality + (1 | Course), data = clean_health,REML=FALSE)

summary(modell)

modelll <- lmer(Depression_Score ~ Stress_Level + Anxiety_Score  + Financial_Stress +
                   (0+CGPA| Course), data = clean_health,REML=FALSE)

summary(modelll)


```


```{r}
plot(model)
qqnorm(residuals(model))
qqline(residuals(model))
qqnorm(residuals(model2))
qqline(residuals(model2))
```

```{r}
r2(model)
r2(model2)
```
```{r}
AIC(model)
AIC(model2)
BIC(model)
BIC(model2)
```
```{r}
ranef(model)
```
```{r}
qqnorm(residuals(model))
qqline(residuals(model))
qqnorm(residuals(model2))
qqline(residuals(model2))
qqnorm(residuals(model3))
qqline(residuals(model3))
```
Here I checked the representation of each course, there is a overrepresentation of medical students, not sure if this is a problem or not.
```{r}
# Count occurrences of each Course
course_counts <- table(clean_health$Course)

print(course_counts)

```
If we are going to follow a neural network some parts of the code above are largely not needed since its a black box approach. Nevertheless we might include it for some insight. My intuition tells me since alot of the features show little variance across students the neural network might need to be rather deep.
As a side note since the depression score is numerical we might want to categorize the values within some ranges so to make it nominal. Or else we can predict a value for a new student, but this is not classification
```{r}
ggplot(clean_health, aes(x = Depression_Score)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "black", alpha = 0.7) +
  ggtitle("Distribution of Depression Scores") +
  xlab("Depression Score") +
  ylab("Frequency") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```
```{r}
depression_distribution_all <- ggplot(clean_health, aes(x = Depression_Score)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  ggtitle("Distribution of Depression Scores") +
  xlab("Depression Score") +
  ylab("Frequency") +
  theme_minimal()

```
```{r}
depression_distribution_all
ggsave("depression_all.png", 
       plot = depression_distribution_all, 
       device = "png", 
       height = 3, width = 6, units = "in")
```

```{r}
depression_counts <- table(clean_health$Depression_Score)
print(depression_counts)

```
```{r}

clean_health$Depression_Group <- ifelse(clean_health$Depression_Score %in% c(0,1,2), "Low (0-2)", "High (3-5)")

# Count occurrences in each group and reorder them
group_counts <- table(clean_health$Depression_Group)
group_counts <- group_counts[c("Low (0-2)", "High (3-5)")]  # Ensuring correct order

# Visualize the counts using a bar plot with y-axis limit and correct order
barplot_heights <- barplot(group_counts, 
                           col = c("lightgreen", "indianred"), 
                           main = "Count of Depression Groups",
                           xlab = "Depression Groups", 
                           ylab = "Count",
                           border = "black",
                           ylim = c(0, 4500))  # Increased limit for space

# Add text labels above bars, adjusting the position slightly lower
text(x = barplot_heights, 
     y = group_counts + 100,  # Slightly adjusting the y position for visibility
     labels = group_counts, 
     pos = 3, 
     cex = 1.2, 
     col = "black")

```
```{r}
# Open a PNG graphics device
png("depression_groups.png", width = 6, height = 4, units = "in", res = 300)

barplot_heights <- barplot(group_counts, 
                           col = c("lightgreen", "indianred"), 
                           main = "Depression Group distribution",
                           xlab = "Depression Groups", 
                           ylab = "Count",
                           border = "black",
                           ylim = c(0, 4500))

text(x = barplot_heights, 
     y = group_counts + 100,  
     labels = group_counts, 
     pos = 3, 
     cex = 1.2, 
     col = "black")

# Close the graphics device to save the file
dev.off()

```

```{r}
# Given counts
low_count <- 3891
high_count <- 3104
total_count <- low_count + high_count

# Accuracy for always predicting the majority class
majority_accuracy <- max(low_count, high_count) / total_count

# Accuracy for a purely random model (50% guessing)
random_accuracy <- (low_count / total_count * 0.5) + (high_count / total_count * 0.5)

# Print results
print(paste("Accuracy of majority class predictor:", round(majority_accuracy * 100, 2), "%"))
print(paste("Accuracy of random guess predictor:", round(random_accuracy * 100, 2), "%"))

```
```{r}
modeler <- lmer(Depression_Score ~ Residence_Type + Substance_Use + CGPA + (1 | Course), data = clean_health)
modeler1 <- lmer(Depression_Score ~ Residence_Type + Substance_Use + (1 | Course), data = clean_health)
summary(model)
```
```{r}
plot(modeler)
qqnorm(residuals(modeler))
qqline(residuals(modeler))
```
```{r}
summary(modeler)
summary(model2)
anova(modeler,modeler1)
```

```{r}
structure(clean_health)
range(clean_health$Semester_Credit_Load)
```

```{r}
#the following plot shows that the Depression Score is getting smaller the higher the GPA becomes, which is unexpected and interesting.
plotnew1 <-ggplot(clean_health, aes(x = CGPA, y = Stress_Level)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(title = "Stress Level vs. CGPA", x = "CGPA", y = "Stress Level") +
  theme_minimal()
```
```{r}
plotnew1
```

```{r}
ggsave(plotnew1, 
       filename = "cgpa_vs_stress.png",
       device = "png",
       height = 3, width = 6, units = "in")
```

```{r}

#Doing the same for each course shows a similar trend, you can also see that the Computer Science students have a significantly larger Depression score
plotnew2 <-ggplot(clean_health, aes(x = CGPA, y = Stress_Level, color = Course)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Strss Level vs. CGPA by Course", x = "CGPA", y = "Stress Level") +
  theme_minimal()
```
```{r}
plotnew2
ggsave(plotnew2, 
       filename = "cgpa_vs_stress_each.png",
       device = "png",
       height = 3, width = 6, units = "in")
```
#On the relation between stress level and depression score, there is slight trend that shows that the stress level has an negative relation with depression score, which I found weird.
```{r}
plotnew3 <-ggplot(clean_health, aes(x = Depression_Score, y = Stress_Level, color = Course)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Depression score vs. Stress level", x = "Depression Score", y = "Stress Level") +
  theme_minimal()
```

```{r}
plotnew3
ggsave(plotnew3, 
       filename = "depression_vs_stress.png",
       device = "png",
       height = 3, width = 6, units = "in")
```

```{r}

#On the following plot you can observe a slight possitive relation between credit load and depression score, although as you can see except from the Computer science students all the others are grouped up very closely together
plotnew4 <-ggplot(clean_health, aes(x = Semester_Credit_Load, y = Stress_Level, color = Course)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Stress level vs. Credit load", x = "Credit load", y = "Stress Level") +
  theme_minimal()
```

```{r}
plotnew4
ggsave(plotnew4, 
       filename = "credit_load_vs_stress_each.png",
       device = "png",
       height = 3, width = 6, units = "in")
```

```{r}
plotnew6 <-ggplot(clean_health, aes(x = Course, y = Stress_Level)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Stress Level by Course", x = "Course", y = "Stress Level") +
  theme_minimal()
```

```{r}
plotnew6
ggsave(plotnew6, 
       filename = "course_stress.png",
       device = "png",
       height = 3, width = 6, units = "in")
```

